Reintroduction of resistant frogs facilitates landscape-scale recovery in the presence of a lethal fungal disease

### 2024-09-28 doc/manuscript/submission_files/translocation.tex
Remove file path for all figures

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% journal abbreviations
\usepackage{jabbrv}
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

The results of the meta-analysis indicate the influence of several
variables on adult survival, and by extension on population viability.
Although we assume that the effects of winter severity and frog sex are
direct effects of the variables themselves, the elevation and donor
population variables are likely serving as proxies for a range of
factors. For example, elevation is correlated with several attributes
that may influence frog survival, including the presence of boulder
habitat (positive correlation) that is important for frog overwintering,
abundance of frog predators including garter snakes (negative
correlation), and frog developmental rates (negative correlation). The
latter can influence the duration of frog life stages that have
differential susceptibility to chytridiomycosis (see below). The
duration of the tadpole and juvenile stages increases with elevation,
but the fact that long-run population growth rate from our viability
model is relatively insensitive to variation in these durations suggests
that developmental rates of early frog life stages have minimal effect
on the predicted long-term viability of reintroduced populations. The
mechanisms underlying the effect of donor population on the probability
of post-translocation frog survival is intriguing, and could indicate
population-specific fitness differences of frogs in the presence of Bd.
Although our study design did not allow us to completely separate the
effects of donor population and recipient site on 1-year adult survival,
we are currently doing so using a modified study design. Specifically,
we are conducting translocations in which frogs from several donor
populations are added concurrently to a single site. Survival of frogs
from each donor population is estimated using CMR methods and the
contribution of each donor population to subsequent generations is
quantified using genetic methods.

@@

For each translocation site, we estimated survival of translocated
frogs, recruitment of new frogs into the adult population, and adult
population size using a site-specific Bayesian open-population
Jolly-Seber CMR model with known additions to the population (i.e.,
translocated cohorts), as implemented by the mrmr
package\textsuperscript{43} and using R Statistical
Software\textsuperscript{56} (v4.4.4) (see \textbf{Supplementary Methods
- Frog population recovery - CMR model structure} for details). Briefly,
the model tracks the states of \emph{M} individuals that comprise a
superpopulation made up of real and
pseudo-individuals\textsuperscript{39}. The possible states of
individuals include ``not recruited'', ``alive'', and ``dead''. The
possible observations of individuals include ``detected'' and ``not
detected''. We assume that individuals that are in the ``not recruited''
or ``dead'' states are never detected (i.e., there are no mistakes in
the individual PIT tag records). We also assume that new recruits were
the result of within-site reproduction and not immigration from adjacent
populations. This assumption is justified by the fact that no \emph{R.
sierrae} populations were present within several kilometers of the
translocation sites. For all models, we used mrmr defaults for priors,
number of chains (4), and warmup and post-warmup iterations (2000 for
each). We evaluated convergence of the Markov chain Monte Carlo (MCMC)
algorithm using trace plots and Gelman-Rubin statistics (Rhat).

@@

In this model we ignore density-dependent recruitment because we were
interested in the growth of the population from an initial
reintroduction and whether this growth was sufficient to prevent
extinction over 50 years following the introduction. We also did not
directly consider the dynamics of Bd in this model. We made this
decision because (i) translocated populations are infected with Bd at
high prevalence\textsuperscript{39}, and (ii) host density does not seem
to play a significant role in multi-year Bd infection dynamics in this
system\textsuperscript{63}. Thus, ignoring Bd infection dynamics and
instead assuming all host vital rates are in the presence of high Bd
prevalence significantly simplifies the model without much loss of
realism. Additional details are provided in \textbf{Supplementary
Methods - Population viability modeling - Incorporating yearly
variability in survival rates} and \textbf{Estimating model parameters}.

@@

\subsubsection{Figures and Legends}\label{figures-and-legends}

@@

\includegraphics[width=0.95\textwidth,height=\textheight]{figures/conceptual_model.png}

@@

\includegraphics[width=0.6\textwidth,height=\textheight]{figures/map_translocation_points.png}

@@

\includegraphics[width=0.6\textwidth,height=\textheight]{figures/translocation_survival_bysiteid.png}

@@

\includegraphics[width=0.8\textwidth,height=\textheight]{figures/cond_effects_plot.png}

@@

\includegraphics[width=0.95\textwidth,height=\textheight]{figures/pop_viability_figures_for_manuscript.png}

>>>>

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

The results of the meta-analysis indicate the influence of several
variables on adult survival, and by extension on population viability.
Although we assume that the effects of winter severity and frog sex are
direct effects of the variables themselves, the elevation and donor
population variables are likely serving as proxies for a range of
factors. For example, elevation is correlated with several attributes
that may influence frog survival, including the presence of boulder
habitat (positive correlation) that is important for frog overwintering,
abundance of frog predators including garter snakes (negative
correlation), and frog developmental rates (negative correlation). The
latter can influence the duration of frog life stages that have
differential susceptibility to chytridiomycosis (see below). The
duration of the tadpole and juvenile stages increases with elevation,
but the fact that long-run population growth rate from our viability
model is relatively insensitive to variation in these durations suggests
that developmental rates of early frog life stages have minimal effect
on the predicted long-term viability of reintroduced populations. The
mechanisms underlying the effect of donor population on the probability
of post-translocation frog survival is intriguing, and could indicate
population-specific fitness differences of frogs in the presence of Bd.
Although our study design did not allow us to completely separate the
effects of donor population and recipient site on 1-year adult survival,
we are currently doing so using a modified study design. Specifically,
we are conducting reintroductions in which frogs from several donor
populations are added concurrently to a single site. Survival of frogs
from each donor population is estimated using CMR methods and the
contribution of each donor population to subsequent generations is
quantified using genetic methods.

@@

For each translocation site, we estimated survival of translocated
frogs, recruitment of new frogs into the adult population, and adult
population size using a site-specific Bayesian open-population
Jolly-Seber CMR model with known additions to the population (i.e.,
translocated cohorts), as implemented by the mrmr
package\textsuperscript{43} and using R Statistical
Software\textsuperscript{56} (v4.4.4) (see \textbf{Supplementary Methods
- Frog population recovery - CMR model structure} for details). Briefly,
the model tracks the states of \emph{M} individuals that comprise a
superpopulation made up of real and
pseudo-individuals\textsuperscript{39}. The possible states of
individuals include ``not recruited'', ``alive'', and ``dead''. The
possible observations of individuals include ``detected'' and ``not
detected''. We assume that individuals that are in the ``not recruited''
or ``dead'' states are never detected (i.e., there are no mistakes in
the individual PIT tag records). We also assume that new recruits were
the result of within-site reproduction and not immigration from adjacent
populations. This assumption is justified by the fact that no \emph{R.
sierrae} populations were present within several kilometers of the
translocation sites. For all models, we used mrmr defaults for priors,
\_number of chains (4), and warmup and post-warmup iterations (2000 for
each). We evaluated convergence of the Markov chain Monte Carlo (MCMC)
algorithm using trace plots and Gelman-Rubin statistics (Rhat).

@@

In this model we ignore density-dependent recruitment because we were
interested in the growth of the population from an initial
reintroduction and whether this growth was sufficient to prevent
extinction over 50 years following the introduction. We also did not
directly consider the dynamics of Bd in this model. We made this
decision because (i) translocated populations are infected with Bd at
high prevalence\textsuperscript{39}, and (ii) host density does not seem
to play a significant role in multi-year Bd infection dynamics in this
system\textsuperscript{63}. Thus, ignoring Bd infection dynamics and
instead assuming all host vital rates are in the presence of high Bd
prevalence significantly simplifies the model without\_ much loss of
realism. Additional details are provided in \textbf{Supplementary
Methods - Population viability modeling - Incorporating yearly
variability in survival rates} and \textbf{Estimating model parameters}.

@@

\subsubsection{Figures and Figure
Legends}\label{figures-and-figure-legends}

@@

\includegraphics{conceptual_model_180mm.pdf}

@@

\includegraphics{map_translocation_points_88mm.pdf}

@@

\includegraphics{translocation_survival_bysiteid_88mm.pdf}

@@

\includegraphics{conditional_effects_plot_88mm.pdf}

@@

\includegraphics{pop_viability_figures_for_manuscript_180mm.pdf}
